---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
## Environment setup
```{r envsetup, echo=TRUE}
suppressMessages(library(dplyr))
library(ggplot2)
setwd("C:/MOOC/module5/assessments1/RepData_PeerAssessment1")
```

## Loading and preprocessing the data
```{r loaddata, echo=TRUE}
data = read.csv("activity.csv",header=T)
data$date = as.Date(data$date,"%Y-%m-%d")
data = tbl_df(data)
```

## What is mean total number of steps taken per day?
```{r dailytotalsteps, echp=TRUE}
daily_steps <- data %>%
                  group_by(date) %>%
                  summarise(total_steps = sum(steps, na.rm=TRUE))

hist(daily_steps$total_steps,
main=" ",
breaks=10,
xlab="Total Number of Steps Taken Daily")
```

```{r dailymeansteps, echo=TRUE}
daily_meansteps <- data %>%
                  group_by(date) %>%
                  summarise(mean_steps = mean(steps, na.rm=TRUE))
daily_meansteps
mean(daily_meansteps$mean_steps,na.rm=TRUE)
```

```{r dailymediansteps, echo=TRUE}
daily_mediansteps <- data %>%
                  filter(!is.na(steps) & steps > 0) %>%
                  group_by(date) %>%
                  summarise(median_steps = median(steps))
daily_mediansteps
median(daily_mediansteps$median_steps)
```
## What is the average daily activity pattern?
```{r averagedailyactivity, echo=TRUE}
avg_daily_pattern <- data %>%
                  filter(!is.na(steps) & steps > 0) %>%
                  group_by(interval) %>%
                  summarise(avg_daily_steps = mean(steps))
avg_daily_pattern %>% arrange(desc(avg_daily_steps))

qplot(x=interval, y=avg_daily_steps, data = avg_daily_pattern,  geom = "line",
      xlab="5-Minute Interval (military time)",
      ylab="Number of Step Count",
      main="Average Number of Steps Taken Averaged Across All Days"
     )

```
  
Observations:
At interval 8:35am with the highest mean steps of 352.48.

## Imputing missing values
```{r totalmissingvalues, echo=TRUE}

daily_nasteps <- data %>%
                  filter(is.na(steps)) %>%
                  group_by(date) %>%
                  summarise(count_nasteps = sum(is.na(steps)))
daily_nasteps
sum(daily_nasteps$count_nasteps)
```

```{r fillmissingvalues, echo=TRUE}
imput_median = median(daily_mediansteps$median_steps)

imput_data = data
imput_data$steps[is.na(imput_data$steps)] <- imput_median
```

```{r histomeanmedian, echo=TRUE}
imput_daily_steps <- imput_data %>%
                  group_by(date) %>%
                  summarise(total_steps = sum(steps))

hist(imput_daily_steps$total_steps,
main=" ",
breaks=10,
xlab="Total Number of Steps Taken Daily")

imput_daily_meansteps <- imput_data %>%
                  group_by(date) %>%
                  summarise(mean_steps = mean(steps))
imput_daily_meansteps
mean(imput_daily_meansteps$mean_steps)

imput_daily_mediansteps <- imput_data %>% 
                  filter(steps > 0) %>%
                  group_by(date) %>%
                  summarise(median_steps = median(steps))
imput_daily_mediansteps
median(imput_daily_mediansteps$median_steps)
```
  
The mean are close to each other while the median has the same value.  
It seem to have little impact.

## Are there differences in activity patterns between weekdays and weekends?
```{r activitypatterns, echo=TRUE}
Sys.setlocale("LC_TIME","English")

dataweek = data %>%
               mutate(daytype = ifelse(weekdays(date) %in% c("Satuday", "Sunday"), 
                               "weekend", "weekday")) %>%
               filter(!is.na(steps) & steps > 0) %>% 
               group_by(interval,daytype) %>%
               summarise(avg_daily_steps = mean(steps))

dataweek

p2 <- qplot(x=interval, y=avg_daily_steps, data = dataweek,  geom = "line",
      xlab="5-Minute Interval (military time)",
      ylab="Number of Step Count",
      main="Average Number of Steps Taken Averaged Across All Days"
      )

p2 + facet_grid(daytype ~ .)
```